On the presentation of orthogonal groups over polynomial rings

### 2019-06-07 dl-paper.tex
Added slicker relations

\subsection{Lemma on the presentation of Steinberg groups}
%Throughout this section $\Phi$ is a root system of type $\rA_{\geq 4}$, $\rD_{\geq 5}$, $\rE_{6,7,8}$ and $A$ is a local ring with maximal ideal $M$.
Let $I\trianglelefteq A$ be an ideal of a commutative ring $A$.
We consider $A[X, X\inv]$ as a graded ring in which $X^{-1}$ has degree $1$.
We denote by $B = B(A, I)$ the subring $A[X^{-1}] + I[X] \subseteq A[X, X\inv]$ with the induced grading.
As an $A$-module $B$ decomposes as $\oplus_{d\in\mathbb Z}B_d$ where $B_d=I \cdot X^d$ for $d<0$, and $B_d=A \cdot X^d$ for $d\geq0$.

We need to introduce notation for certain relations of $\St(\Phi, B)$.
\begin{align}
&\,\,x_{\alpha}(a)x_{\alpha}(b) =  x_{\alpha}(a+b),& a,b\in B_d, \tag{R$1_d$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= x_{\alpha+\beta}(N_{\alpha\beta} \cdot ab),& a\in B_d,\ b\in B_e, \tag{R$2_{d,e}$} \\
&[x_{\alpha}(a),\,x_{\gamma}(b)]= 1,& \alpha \perp \beta,\ a\in B_d,\ b\in B_e, \tag{R$3^\bot_{d,e}$} \\
&[x_{\alpha}(a),\,x_{\beta}(b)]= 1,& \alpha - \beta \in \Phi,\ a\in B_d,\ b\in B_e. \tag{R$3^\angle_{d,e}$}
\end{align}
%For such a ring $R$ one can assume in the definition of the Steinberg group that the generators $x_\alpha(a)$ and relations R1--R3 are defined only for homogeneous elements.

@@

\begin{comment}
\begin{rem}
We can take $\mathbb R^l$ with a base $\chi_1,\ldots,\chi_l$ and visualise root systems
\[\rD_l=\{\pm\chi_i\pm\chi_j\mid1\leq i\neq j\leq l\}\]
and $\rC_l=\rD_l\cup\{\pm2\chi_i\mid1\leq i\leq l\}$. Then we immediately see that $\alpha+\beta\in\rC_l\setminus\rD_l$ exactly means that either $\{\alpha,\beta\}$ or $\{-\alpha,-\beta\}$ equals $\{\chi_i-\chi_j,\chi_i+\chi_j\}$. Then for some $\sigma$ from the Weyl group of $\rD_l$ we get 
\[ \{\sigma\alpha,\sigma\beta\}=\{\chi_{l-1}-\chi_l, \chi_{l-1}+\chi_l\}=\{\alpha_{l-1},\alpha_l\} \]
(the horns of the Dynkin diagram $\rD_l$). Another condition on $\alpha$ and $\beta$ equivalent to $\alpha+\beta\in\rC_l\setminus\rD_l$ is that they do not lie in any subsystem of $\rD_l$ of type $\rA_m$ for $m\geq4$.
\end{rem}

Everywhere below we set $t = X^{-1}$.
\begin{lemma} \label{claim2} In $\St^{\leq n}(\Phi, B)$ relations $\text{R3}^\angle_{d, e}$ imply $\text{R3}^{\bot,0}_{d, e}$ for $1\leq d\leq e \leq n$. \end{lemma}
\begin{proof}
Without loss of generality we may assume that $\alpha, \gamma$ fit into some $\mathsf{A}_3$-triple $(\alpha, \beta, \gamma)$.
 \begin{align*}
 {}^{x_\gamma(t^d)}x_\alpha(bt^{e}) = {}^{x_\gamma(t^d)}[x_{-\beta}(b't^{e}), x_{\alpha+\beta}(1)] & \text{ by $\text{R2}_{e,0}$}\\
 = [x_{-\beta}(b't^{e}), {}^{x_\gamma(t^d)}x_{\alpha+\beta}(1)] & \text{ by $\text{R3}^\angle_{d, e}$}\\
 = [[x_{-\beta-\gamma}(b''t^{e-d}), x_\gamma(t^d)], {}^{x_\gamma(t^d)}x_{\alpha+\beta}(1)] & \text{ by $\text{R2}_{e-d,d}$}\\
 = [x_{-\beta-\gamma}(b''t^{e-d)}), [x_\gamma(t^d), x_{\alpha+\beta}(1)]] & \text{ by~\eqref{eq:H1iii} and $\text{R3}^\bot_{e-d,0}$}\\
 = x_{\alpha}(b'''t^{e}) & \text{ by $\text{R2}_{0,d}$ and $\text{R2}_{e-d,d}$.} \end{align*}
Usual identities for structure constants imply (cf.~\cite[p.~12]{Re75}):
 $b'''=N_{-\beta-\gamma, \alpha+\beta+\gamma} \cdot N_{\gamma, \alpha+\beta} \cdot N_{-\beta-\gamma, \gamma} \cdot N_{-\beta, \alpha+\beta} \cdot b = b,$ from which the claim follows.
\end{proof}

\begin{lemma} \label{claim3}
 In $\St^{\leq n}(\Phi, B)$ relations $\bot_0(d, e)$, $\angle(d, d)$ and $\angle(e, e)$ imply $\bot(d, e)$ for $d\leq e \leq n$.
\end{lemma}
\begin{proof}
 \begin{align*} 
 x_{\alpha+\beta+\gamma}(abt^{e}) = {}^{x_{-\beta}(t^d)}\!x_{\alpha+\beta+\gamma}(abt^{e}) & \text{ by $\text{R3}^{\bot, 0}_{d, e}$}\\
 = {}^{x_{-\beta}(t^d)}\![x_{\alpha+\beta}(b't^{e-d}), x_\gamma(at^d)] & \text{ by $\text{R2}_{e-d,d}$} \\
 = [x_{\alpha}(b't^{e}) \cdot x_{\alpha+\beta}(b't^{e-d}), x_\gamma(at^d)] & \text{ by $\text{R2}_{e-d,d}$ and $\text{R3}^\angle_{d, d}$ }\\
 = {}^{x_{\alpha}(b't^{e})}\!x_{\alpha+\beta+\gamma}(abt^{e}) [x_{\alpha}(b't^{e}), x_\gamma(at^d)] & \text{ by~\eqref{eq:H1ii} and $\text{R2}_{e-d,d}$} \\ = x_{\alpha+\beta+\gamma}(abt^{e}) [x_{\alpha}(b't^{e}), x_\gamma(at^d)] & \text{ by $\text{R3}^\angle_{e, e}$. \qedhere}
\end{align*}
\end{proof}
\end{comment}

\begin{align}
 \label{eq:H1ii}  [ab, c] = {}^a[b, c] \cdot [a,c];&\\ %= [a,[b,c]]\cdot [b,c] \cdot [a,c];&\\
 \label{eq:H1iii} [a,c]   = 1    \text{ implies } [a, [b,c]] = [[a,b],{}^bc].&
\end{align}

\begin{align*} 
 [x_\alpha(\xi), x_{\gamma}(bt^{e})] = [x_\alpha(\xi), [x_{\beta+\gamma}(b't^{e-1}), x_{-\beta}(t)]] & \text{ by~$\text{R2}_{e-1,1}$} \\
 = [[x_\alpha(\xi), x_{\beta+\gamma}(b't^{e-1})], {}^{x_{\beta+\gamma}(b't^{e-1})}\!x_{-\beta}(t)] & \text{ by \eqref{eq:H1iii} and $\text{R3}^\angle_{0,1}$} \\
 = {}^{x_{\beta+\gamma}(b't^{e-1})}\![x_{\alpha+\beta+\gamma}(b't^{e-1}\xi), x_{-\beta}(t)] & \text{ by $\text{R2}_{0,e-1}$ and $\text{R3}^\angle_{e-1,e-1}$} \\
 = 1 & \text{ by $\text{R3}^\bot_{e-1,1}$.}
\end{align*}

@@

\begin{align*}
[x_{il}(abc), x_{jk}(b)] &= 1 & \text{R3}_{n,m} \\
[x_{ij}(a), x_{kl}(c)] &= 1 & \text{R3}_{n-m, 0}\\
[x_{ik}(ab), x_{kl}(c)] &= x_{il}(abc), & \text{R2}_{n, 0} \\
[x_{jk}(b), x_{kl}(c)] &= x_{jl}(bc), & \text{R2}_{m, 0}\\
[x_{ij}(a), x_{jk}(b)] &= x_{ik}(ab), & \text{R2}_{n-m, m}\\
[x_{ij}(a), x_{jl}(bc)] &= x_{il}(abc),& \text{R2}_{n-m, m}\\
[x_{jk}(b), x_{kl}(c)] &= x_{jl}(bc), & \text{R2}_{m, 0}\\
[x_{ij}(a), x_{jk}(b)] &= x_{ik}(ab) & \text{R2}_{n-m, m}
\end{align*}
\end{comment}

@@

\begin{comment}
\begin{align*} {}^{x_\gamma(X^{-1})}x_\alpha(bX^{-1}) = {}^{x_\gamma(X^{-1})}[x_{-\beta}(b'X^{-1}), x_{\alpha+\beta}(1)] & \text{ by R$2_1$}\\
 = [x_{-\beta}(b'X^{-1}), {}^{x_\gamma(X^{-1})}x_{\alpha+\beta}(1)] & \text{ by R$3_1$ for sharp angles}\\
 = [[x_{-\beta-\gamma}(b''), x_\gamma(X^{-1})], {}^{x_\gamma(X^{-1})}x_{\alpha+\beta}(1)] & \text{ by R$2_1$}\\
 = [x_{-\beta-\gamma}(b''), [x_\gamma(X^{-1}), x_{\alpha+\beta}(1)]] & \text{ by R$3_0$}\\
 = x_{\alpha}(b'''X^{-1}) & \text{ by R$2_1$.} \end{align*}

\begin{align*} x_{\alpha+\beta+\gamma}(abX^{-1}) = {}^{x_{-\beta}(X^{-1})}\!x_{\alpha+\beta+\gamma}(abX^{-1}) & \text{ by the above relation}\\
 = {}^{x_{-\beta}(X^{-1})}\![x_{\alpha+\beta}(b'), x_\gamma(aX^{-1})] & \text{ by R$2_1$} \\
 = [x_{\alpha}(b'X^{-1}) x_{\alpha+\beta}(b'), x_\gamma(aX^{-1})] & \text{ by R$2_1$ and R$3_1$ for sharp angles }\\
 = {}^{x_{\alpha}(b'X^{-1})}\!x_{\alpha+\beta+\gamma}(abX^{-1}) [x_{\alpha}(b'X^{-1}), x_\gamma(aX^{-1})] & \text{ by R$2_1$} \\
 = x_{\alpha+\beta+\gamma}(abX^{-1}) [x_{\alpha}(b'X^{-1}), x_\gamma(aX^{-1})] & \text{ by R$3_1$ for sharp angles.} \end{align*}
\end{comment}

\include{rehmann}

>>>>

\subsection{Lemma on the presentation of Steinberg groups}
%Throughout this section $\Phi$ is a root system of type $\rA_{\geq 4}$, $\rD_{\geq 5}$, $\rE_{6,7,8}$ and $A$ is a local ring with maximal ideal $M$.
Let $I\trianglelefteq A$ be an ideal of a commutative ring $A$.
We consider $A[X, X\inv]$ as a $\mathbb{Z}$-graded ring in which $X^{-1}$ has degree $1$.
We denote by $B = B(A, I)$ the subring $A[X^{-1}] + I[X] \subseteq A[X, X\inv]$ with the induced grading.
As an $A$-module $B$ decomposes as $\oplus_{d\in\mathbb Z}B_d$ where $B_d=I \cdot X^d$ for $d>0$, and $B_d=A \cdot X^d$ for $d\leq0$.
Obviously, $B = A[X^{-1}]$ in the case $I=0$ and $B = A[X, X^{-1}]$ in the case $I=A$.

We need to introduce notation for certain relations of $\St(\Phi, B)$. 
\begin{align}
&\,\,x_{\alpha}(a)x_{\alpha}(a') =  x_{\alpha}(a+a'),& \tag{R$1_d$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= x_{\alpha+\beta}(N_{\alpha, \beta} \cdot ab),& \tag{R$2_{d,e}$} \\
&[x_{\alpha}(a),\,x_{\gamma}(b)]= 1,& \alpha \perp \beta,\ \tag{R$3^\bot_{d,e}$} \\
&[x_{\alpha}(a),\,x_{\beta}(b)]= 1,& \alpha - \beta \in \Phi,\ \tag{R$3^\angle_{d,e}$}
\end{align}
where $a, a' \in B_d$, $b\in R_e$. 
%For such a ring $R$ one can assume in the definition of the Steinberg group that the generators $x_\alpha(a)$ and relations R1--R3 are defined only for homogeneous elements.

@@

\begin{comment}
\begin{rem}
We can take $\mathbb R^l$ with a base $\chi_1,\ldots,\chi_l$ and visualise root systems
\[\rD_l=\{\pm\chi_i\pm\chi_j\mid1\leq i\neq j\leq l\}\]
and $\rC_l=\rD_l\cup\{\pm2\chi_i\mid1\leq i\leq l\}$. Then we immediately see that $\alpha+\beta\in\rC_l\setminus\rD_l$ exactly means that either $\{\alpha,\beta\}$ or $\{-\alpha,-\beta\}$ equals $\{\chi_i-\chi_j,\chi_i+\chi_j\}$. Then for some $\sigma$ from the Weyl group of $\rD_l$ we get 
\[ \{\sigma\alpha,\sigma\beta\}=\{\chi_{l-1}-\chi_l, \chi_{l-1}+\chi_l\}=\{\alpha_{l-1},\alpha_l\} \]
(the horns of the Dynkin diagram $\rD_l$). Another condition on $\alpha$ and $\beta$ equivalent to $\alpha+\beta\in\rC_l\setminus\rD_l$ is that they do not lie in any subsystem of $\rD_l$ of type $\rA_m$ for $m\geq4$.
\end{rem}
Everywhere below we set $t = X^{-1}$.
\begin{lemma} \label{claim2} In $\St^{\leq n}(\Phi, B)$ relations $\text{R3}^\angle_{d, e}$ imply $\text{R3}^{\bot,0}_{d, e}$ for $1\leq d\leq e \leq n$. \end{lemma}
\begin{proof}
Without loss of generality we may assume that $\alpha, \gamma$ fit into some $\mathsf{A}_3$-triple $(\alpha, \beta, \gamma)$.
 \begin{align*}
 {}^{x_\gamma(t^d)}x_\alpha(bt^{e}) = {}^{x_\gamma(t^d)}[x_{-\beta}(b't^{e}), x_{\alpha+\beta}(1)] & \text{ by $\text{R2}_{e,0}$}\\
 = [x_{-\beta}(b't^{e}), {}^{x_\gamma(t^d)}x_{\alpha+\beta}(1)] & \text{ by $\text{R3}^\angle_{d, e}$}\\
 = [[x_{-\beta-\gamma}(b''t^{e-d}), x_\gamma(t^d)], {}^{x_\gamma(t^d)}x_{\alpha+\beta}(1)] & \text{ by $\text{R2}_{e-d,d}$}\\
 = [x_{-\beta-\gamma}(b''t^{e-d)}), [x_\gamma(t^d), x_{\alpha+\beta}(1)]] & \text{ by~\eqref{eq:H1iii} and $\text{R3}^\bot_{e-d,0}$}\\
 = x_{\alpha}(b'''t^{e}) & \text{ by $\text{R2}_{0,d}$ and $\text{R2}_{e-d,d}$.} \end{align*}
Usual identities for structure constants imply (cf.~\cite[p.~12]{Re75}):
 $b'''=N_{-\beta-\gamma, \alpha+\beta+\gamma} \cdot N_{\gamma, \alpha+\beta} \cdot N_{-\beta-\gamma, \gamma} \cdot N_{-\beta, \alpha+\beta} \cdot b = b,$ from which the claim follows.
\end{proof}

\begin{lemma} \label{claim3}
 In $\St^{\leq n}(\Phi, B)$ relations $\bot_0(d, e)$, $\angle(d, d)$ and $\angle(e, e)$ imply $\bot(d, e)$ for $d\leq e \leq n$.
\end{lemma}
\begin{proof}
 \begin{align*} 
 x_{\alpha+\beta+\gamma}(abt^{e}) = {}^{x_{-\beta}(t^d)}\!x_{\alpha+\beta+\gamma}(abt^{e}) & \text{ by $\text{R3}^{\bot, 0}_{d, e}$}\\
 = {}^{x_{-\beta}(t^d)}\![x_{\alpha+\beta}(b't^{e-d}), x_\gamma(at^d)] & \text{ by $\text{R2}_{e-d,d}$} \\
 = [x_{\alpha}(b't^{e}) \cdot x_{\alpha+\beta}(b't^{e-d}), x_\gamma(at^d)] & \text{ by $\text{R2}_{e-d,d}$ and $\text{R3}^\angle_{d, d}$ }\\
 = {}^{x_{\alpha}(b't^{e})}\!x_{\alpha+\beta+\gamma}(abt^{e}) [x_{\alpha}(b't^{e}), x_\gamma(at^d)] & \text{ by~\eqref{eq:H1ii} and $\text{R2}_{e-d,d}$} \\ = x_{\alpha+\beta+\gamma}(abt^{e}) [x_{\alpha}(b't^{e}), x_\gamma(at^d)] & \text{ by $\text{R3}^\angle_{e, e}$. \qedhere}
\end{align*}
\end{proof}
\end{comment}
\begin{align}
 \label{eq:H1ii}  [ab, c] = {}^a[b, c] \cdot [a,c];&\\ %= [a,[b,c]]\cdot [b,c] \cdot [a,c];&\\
 \label{eq:H1iii} [a,c]   = 1    \text{ implies } [a, [b,c]] = [[a,b],{}^bc].&
\end{align}

In the case $1\leq d \leq e \leq n$ the assertion follows from the following computation:
\begin{align*} x_{\alpha+\beta+\gamma}(abt^e) = [x_{\alpha+\beta}(at^{e-d}), [x_{-\beta}(t^d), x_{\beta+\gamma}(b)] ] & \text{ by $\text{R2}_{d,0}$, $\text{R2}_{d, e-d}$  } \\ 
 = [[x_{\alpha+\beta}(at^{e-d}), x_{-\beta}(t^d)], {}^{x_{-\beta}(t^d)} x_{\beta+\gamma}(b)] & \text{ by~\eqref{eq:H1iii}, $\text{R3}^\bot_{0,e-d}$} \\ 
 = {}^{x_{-\beta}(t^d)} [x_{\alpha}(at^e), x_{\beta+\gamma}(b)] & \text{ by $\text{R2}_{e-d,d}$, $\text{R3}^\angle_{d,e}$} \\ 
 = {}^{x_{-\beta}(t^d)} [x_{\alpha + \beta}(at^{e-d}), x_{\gamma}(bt^d)] & \text{ by $\text{R2}_{e,0}$, $\text{R2}_{e-d,d}$} \\ 
 = [x_{\alpha}(at^e) \cdot x_{\alpha + \beta}(at^{e-d}), x_{\gamma}(bt^d)] & \text{ by $\text{R2}_{e-d,d}$, $\text{R3}^\angle_{d,d}$} \\ 
 = {}^{x_\alpha(at^e)}x_{\alpha+\beta+\gamma}(abt^e) \cdot [x_\alpha(at^e), x_\gamma(bt^d)] & \text{ by~\eqref{eq:H1ii}, $\text{R2}_{e-d,d}$} \\
 = x_{\alpha+\beta+\gamma}(abt^e) \cdot [x_\alpha(at^e), x_\gamma(bt^d)] & \text{ by $\text{R3}^\angle_{e,e}$}. \end{align*}

In the case $d < 1\leq e \leq n$
\begin{align*} 
 [x_\alpha(\xi), x_{\gamma}(bt^{e})] = [x_\alpha(\xi), [x_{\beta+\gamma}(b't^{e-1}), x_{-\beta}(t)]] & \text{ by~$\text{R2}_{e-1,1}$ } \\
 = [[x_\alpha(\xi), x_{\beta+\gamma}(b't^{e-1})], {}^{x_{\beta+\gamma}(b't^{e-1})}\!x_{-\beta}(t)] & \text{ by \eqref{eq:H1iii} and $\text{R3}^\angle_{d,1}$ } \\
 = {}^{x_{\beta+\gamma}(b't^{e-1})}\![x_{\alpha+\beta+\gamma}(b't^{e-1}\xi), x_{-\beta}(t)] & \text{ by $\text{R2}_{d,e-1}$ and $\text{R3}^\angle_{e-1,e-1}$  } \\
 = 1 & \text{ by $\text{R3}^\bot_{d+e-1,1}$.} \end{align*}

@@

\begin{align*}
[x_{il}(abc), x_{jk}(b)] &= 1 & \text{R3}_{n,m} \\
[x_{ij}(a), x_{kl}(c)] &= 1 & \text{R3}_{n-m, 0}\\
[x_{ik}(ab), x_{kl}(c)] &= x_{il}(abc), & \text{R2}_{n, 0} \\
[x_{jk}(b), x_{kl}(c)] &= x_{jl}(bc), & \text{R2}_{m, 0}\\
[x_{ij}(a), x_{jk}(b)] &= x_{ik}(ab), & \text{R2}_{n-m, m}\\
[x_{ij}(a), x_{jl}(bc)] &= x_{il}(abc),& \text{R2}_{n-m, m}\\
[x_{jk}(b), x_{kl}(c)] &= x_{jl}(bc), & \text{R2}_{m, 0}\\
[x_{ij}(a), x_{jk}(b)] &= x_{ik}(ab) & \text{R2}_{n-m, m}
\end{align*}

@@

\begin{align*} {}^{x_\gamma(X^{-1})}x_\alpha(bX^{-1}) = {}^{x_\gamma(X^{-1})}[x_{-\beta}(b'X^{-1}), x_{\alpha+\beta}(1)] & \text{ by R$2_1$}\\
 = [x_{-\beta}(b'X^{-1}), {}^{x_\gamma(X^{-1})}x_{\alpha+\beta}(1)] & \text{ by R$3_1$ for sharp angles}\\
 = [[x_{-\beta-\gamma}(b''), x_\gamma(X^{-1})], {}^{x_\gamma(X^{-1})}x_{\alpha+\beta}(1)] & \text{ by R$2_1$}\\
 = [x_{-\beta-\gamma}(b''), [x_\gamma(X^{-1}), x_{\alpha+\beta}(1)]] & \text{ by R$3_0$}\\
 = x_{\alpha}(b'''X^{-1}) & \text{ by R$2_1$.} \end{align*}

\begin{align*} x_{\alpha+\beta+\gamma}(abX^{-1}) = {}^{x_{-\beta}(X^{-1})}\!x_{\alpha+\beta+\gamma}(abX^{-1}) & \text{ by the above relation}\\
 = {}^{x_{-\beta}(X^{-1})}\![x_{\alpha+\beta}(b'), x_\gamma(aX^{-1})] & \text{ by R$2_1$} \\
 = [x_{\alpha}(b'X^{-1}) x_{\alpha+\beta}(b'), x_\gamma(aX^{-1})] & \text{ by R$2_1$ and R$3_1$ for sharp angles }\\
 = {}^{x_{\alpha}(b'X^{-1})}\!x_{\alpha+\beta+\gamma}(abX^{-1}) [x_{\alpha}(b'X^{-1}), x_\gamma(aX^{-1})] & \text{ by R$2_1$} \\
 = x_{\alpha+\beta+\gamma}(abX^{-1}) [x_{\alpha}(b'X^{-1}), x_\gamma(aX^{-1})] & \text{ by R$3_1$ for sharp angles.} \end{align*}

\include{rehmann}
\end{comment}
